Transfer learning for \'echelle spectroscopy with \texttt{blas\'e}

### 2022-03-15 paper/paper1/ms.tex
Add appendix describing logarithmic alternative

At this stage, we have the choice of whether to work in linear or log scale flux units. Adopting the log of the flux would ensure that the cloned model possesses only positive flux values, a desirable trait of any physical spectral model. We chose instead to stick with native linear flux scaling because of the availability of convenient approximations to astrophysically relevant line profiles in the linear scaling domain. We normalize the spectrum by its median value since we do not intend to apply these methods to absolutely flux calibrated spectra.

@@

\section{Autodiff themes} \label{appendix:tools}

Here are some more details about autodiff

>>>>

At this stage, we have the choice of whether to work in linear or log scale flux units. Adopting the log of the flux would ensure that the cloned model possesses only positive flux values, a desirable trait of any physical spectral model. We have implemented both modes in \texttt{blas\'e}, allowing users to choose their preference.  We only narrate the linear flux unit description in the main text of this document, for the sake of clarity, and since most practitioners tend to think of flux in terms of linear flux units.  The data-model comparison step will always take place in linear flux units, so the only operational difference is the behavior for deep and saturated lines.  Appendix \censorbox{XX} lists the equations adjusted to log flux units.

@@

\section{Log flux scaling mode} \label{appendixLogScale}

Here we illustrate how \texttt{blas'e} gets altered when applying the logarithmic flux pre-processing step.  First, we compute the natural log of the flux directly on the precomputed synthetic spectrum in its absolute flux scaling and native pixel sampling:

\begin{eqnarray}
    \ln{\mathsf{S}} = \ln{\mathsf{S}_{\rm abs}} - \ln{\mathsf{B}} - \mathsf{P}
    \label{eqnlogFlat}
\end{eqnarray}

We simply ``rebrand'' $\mathsf{P}$ as residing in logarithmic flux units, and disregard it since it is largely a nuisance parameter anyways.  We then treat the \texttt{blas\'e} clone model as a sum of opacities, retaining the Voigt profile:

\begin{eqnarray}
    \ln{\mathsf{S}_{\rm clone}} = -\sum_{j=1}^{N_{\mathrm{lines}}} a_j \mathsf{V}_j \label{equationOpacitySum}
\end{eqnarray}

Here, the $a_j$'s have also been slightly rebranded from their meaning in Equation \ref{equation1}.  We still want to enforce only absorption lines---and not spurious emission lines---so we use the sample trick of sampling the $a_j$'s in log and then exponentiating them to get guaranteed positive values.  Note that Equations \ref{equation1} and \ref{equationOpacitySum} carry modified meanings for the Voigt profile.  Specifically, Equation \ref{equation1} can be viewed as the Taylor Series expansion for \ref{equationOpacitySum} in the limit of small opacities:

\begin{eqnarray}
    e^{-a_j \mathsf{V}_j} \approx (1-a_j\mathsf{V}_j) \label{eqnTaylor}
\end{eqnarray}

Both equations are approximate. A real stellar atmosphere's lineshape arises from a sum of disparate Voigt profiles weighted along a nonuniform column of gas, whereas here we have assumed the column of gas is approximated as a single uniform isothermal backlit layer.  A sum of unlike-Voigt profiles is not exactly equal to any single Voigt profile.  Theoreticians may resonate with this more ``first principles'' representation, while data practitioners may find Equation \ref{equation1} more natural, so to some extent the choice is a matter of taste.

The sparse matrix gets rebranded as filled with opacity values, instead of log-fluxes, but operationally remains the same. All subsequent steps operate on the summed-and-exponentiated opacities, behaving identically to their linear counterparts.  For example, we exponentiate before computing the residuals and data-model comparison, $\mathsf{R} = e^{\ln{\mathsf{S}}} - e^{\ln{\mathsf{S}_{\rm clone}}}$.
